Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-06-07 paper/paper.tex
Added bibliography
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@619 3fc257bb-8550-0410-81b3-a5c103da83f7
More.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@625 3fc257bb-8550-0410-81b3-a5c103da83f7
Lots of small edits + some new XMT results.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@628 3fc257bb-8550-0410-81b3-a5c103da83f7

\documentclass[11pt]{article}
\usepackage[pdftex]{graphicx}

@@

\author{
Steven J.~Plimpton, Karen D.~Devine, Others? \\
Sandia National Laboratories \\
Albuquerque, NM \\
sjplimp@sandia.gov
}

@@

We describe a parallel library written with message-passing (MPI)
calls that allows algorithms to be expressed in the MapReduce
paradigm.  This means the calling program does not need to include
explicit parallel code, but instead provides ``map'' and ``reduce''
functions which operate independently on elements of a data set
distributed across processors.  The library performs needed data
movement between processors.  We describe how typical MapReduce
funtionality can be implemented in an MPI context, and also in an
out-of-core manner for data sets that do not fit within the aggregate
memory of a parallel machine.  Our motivation for creating this
library was to enable graph algorithms to be written as MapReduce
operations, allowing processing of Terabyte-scale data sets.  We
outline MapReduce versions of several such algorithms: vertex ranking
via PageRank, triangle finding, connected component identification,
Luby's algorithm for maximally independent sets, and single-source
shortest-path calculation.  To test the algorithms on arbitrarily
large artificial graphs we generate randomized R-MAT matrices in
parallel; a MapReduce version of this operation is also described.
Performance and scalability results for the various algorithms are
presented for varying size graphs on a distributed-memory cluster.
For some cases, we compare the results with non-MapReduce algorithms,
different machines, and different MapReduce software, namely Hadoop.
Our open-source library is written in C++, is callable from C++, C,
Fortran, or scripting languages such as Python, and can run on any
parallel platform that supports MPI.

@@

\bibliographystyle{aip}
\bibliography{paper}

>>>>

\documentclass[11pt]{article}
\usepackage[pdftex]{graphicx}
\usepackage{url}

@@

\author{
Steven J.~Plimpton and Karen D.~Devine \\
Sandia National Laboratories \\
Albuquerque, NM \\
sjplimp@sandia.gov
}

@@

We describe a parallel library written with message-passing (MPI)
calls that allows algorithms to be expressed in the MapReduce
paradigm.  This means the calling program does not need to include
explicit parallel code, but instead provides ``map'' and ``reduce''
functions that operate independently on elements of a data set
distributed across processors.  The library performs needed data
movement between processors.  We describe how typical MapReduce
funtionality can be implemented in an MPI context, and also in an
out-of-core manner for data sets that do not fit within the aggregate
memory of a parallel machine.  Our motivation for creating this
library was to enable graph algorithms to be written as MapReduce
operations, allowing processing of Terabyte-scale data sets.  We
outline MapReduce versions of several such algorithms: vertex ranking
via PageRank, triangle finding, connected component identification,
Luby's algorithm for maximally independent sets, and single-source
shortest-path calculation.  To test the algorithms on arbitrarily
large artificial graphs we generate randomized R-MAT matrices in
parallel; a MapReduce version of this operation is also described.
Performance and scalability results for the various algorithms are
presented for varying size graphs on a distributed-memory cluster.
For some cases, we compare the results with non-MapReduce algorithms,
different machines, and different MapReduce software, namely Hadoop.
Our open-source library is written in C++, is callable from C++, C,
Fortran, or scripting languages such as Python, and can run on any
parallel platform that supports MPI.

@@

\bibliographystyle{abbrv}
\bibliography{paper}

### 2010-06-07 paper/results.tex
Lots of small edits + some new XMT results.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@628 3fc257bb-8550-0410-81b3-a5c103da83f7

We ran each of the algorithms on four R-MAT graphs of
different sizes, each on a varying number of processors.  Details of the
input data are shown in Table~\ref{t:rmats}.
The {\it small} problem size (around 8M edges) can typically be used
on a single processor without significant out-of-core operations.
The {\it medium} problem size (around 134M edges) is a 
problem that can be run
on a single processor with out-of-core operations; larger processor 
configurations, however, do not necessarily need out-of-core operations.
The {\it large} problem size (around 2B
edges) requires out-of-core operations and higher processor counts.
%The {\it x-large} problem size
%(around 34B edges) requires most of the machine to run.
All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19, 0.05)$.

@@

The resulting timings give a sense of the inherent scalability of the
MapReduce algorithms as graph size grows on a fixed number of
processors, and of the parallel scalability for computing on a graph
of fixed size on a growing number of processors.  Where available, we
compare the MapReduce algorithm with other parallel implementations.
We compute parallel efficiency on $p$ processors
as $({time}_{m} \times m) / ({time}_p \times p)$
where $m$ is the smallest number of processors on which the experiment
was run, and ${time}_m$ is the execution time required on $m$ processors.

@@

In Figure~\ref{f:rmat}, we show the scalability of the R-MAT generation 
algorithm (Figure~\ref{fig:rmat2}) for RMAT-20, RMAT-24 and RMAT-28.  For
RMAT-20 and RMAT-24, superlinear speed-up is shown.  This speed-up is due
to the decreased amount of file I/O needed with greater numbers of 
processors; with a larger total memory, a fixed-size problem requires
less I/O with more processors.  For RMAT-28, where most operation require
significant out-of-core operation, parallel efficiency ranges from 52\% to
97\%.  Comparisons between the enhanced algorithm (Figure~\ref{fig:rmat2}) and
the original algorithm (Figure~\ref{fig:rmat}) showed approximately 
10\% reduction in execution time for the enhanced algorithm.

KDD NOTE:  Add XMT or PBGL results here.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_rmat.pdf}
\caption{Performance of R-MAT generation algorithm (Figure~\ref{fig:rmat2}).}
\label{f:rmat}
\end{figure}

\subsection{PageRank results}

In Figure~\ref{f:pr}, we show the performance of the MR-MPI PageRank algorithm
(Figure~\ref{fig:pr2}) compared to a
distributed-memory matrix-based implementation using the linear
algebra toolkit Trilinos~\cite{Trilinos-Overview} 
and a multi-threaded
implementation
in the Multi-Threaded Graph Library (MTGL)~\cite{MTGL}.
The matrix-based distributed-memory implementation of PageRank
uses Trilinos' Epetra matrix/vector classes to represent the graph and
PageRank vector.
Rows of matrix $A$ and the associated entries of the PageRank vector $x$
are uniquely assigned to processors; a random permutation of the input
matrix achieves processor load balance.
Interprocessor communication gathers $x$ values for matrix-vector
multiplication and sums partial products into the $y$ vector.
Most communication is point-to-point communication,
but some global communication is needed for computing
residuals and norms of $x$ and $y$.
In the multi-threaded MTGL~\cite{MTGL} implementation,
rank propagates via adjacency list traversal
in a compressed sparse-row data structure.
To maintain scalability, code must
be written so that a single thread spawns the loop that processes all
in-neighbors of a given vertex; this detail enables the compiler to generate
hotspot-free code.

@@

KDD NOTE:  Need to add XMT results; PBGL would be nice, too.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_pagerank.pdf}
\caption{Comparison of PageRank implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{f:pr}
\end{figure}

@@

In Figure~\ref{f:tri}, we show the performance of the triangle finding 
algorithm (Figure~\ref{fig:tri}).  Execution times for this algorithm were
too large to allow the problem to be run with RMAT-28 on our cluster.
Parallel efficiencies for RMAT-20 ranged from 80\% to 140\%; for RMAT-24, they
ranged from 50\% to 120\%.

KDD NOTE:  Need to add XMT results; triangle enumeration is not implemented in PBGL.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_tri.pdf}
\caption{Performance of the MR-MPI triangle-finding algorithm (Figure~\ref{fig:tri}).}
\label{f:tri}
\end{figure}

@@

KDD NOTE:  Need to add XMT results.

@@

The execution times for the maximally independent set algorithm
(Figure~\ref{fig:luby}) are shown in Figure~\ref{f:luby}.  Like the R-MAT
generation results, superlinear speed-up of the algorithm occurs for RMAT-20
and RMAT-24, as more of the graph fits into processor memory and less file
I/O is needed.  For RMAT-28, the algorithm requires significant out-of-core
operations. In this case, parallel efficiency is nearly perfect going from 
8 to 32 processors; some loss of efficiency is seen at 64 processors,
where the parallel efficiency compared to the 8-processor case is 69\%.

KDD NOTE:  I think I had a bad node for the 64-processor run; I am 
rerunning it now and soon will have revised RMAT-28 results on 64 processors.

KDD NOTE:  Need to add XMT results; PBGL would be nice, too.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_luby.pdf}
\caption{Performance of maximally independent set algorithm (Figure~\ref{fig:luby}).}
\label{f:luby}
\end{figure}

@@

KDD NOTE:  Need to add XMT results; PBGL would be nice, too.

To compare our MR-MPI implementation with a wider set of implementations,
we performed experiments MR-MPI, Hadoop, PBGL (Parallel
Boost Graph Library)~\cite{PBGL} and a multi-threaded implementation on the
Cray XMT using two web graphs: {WebGraphA} with 13.2M
vertices and 31.9M edges, and {WebGraphB} with 187.6M vertices and 531.9M edges.
In Figure~\ref{f:ssspA}, we show execution times for the SSSP algorithm
using {WebGraphA}.   Like our MR-MPI implementation, the Hadoop 
implementation is a Bellman-Ford-style~\cite{Bellman58,Ford62}  algorithm.  
The XMT and PBGL implementations are based on 
delta-stepping~\cite{MeyerSanders98}, and do not require full iterations
over the entire edge list to advance a breadth-first search.  
We observe that the MR-MPI
implementation runs in less time than the Hadoop implementation,but requires
significantly more time than the XMT and PBGL implementations.  
In experiments with {WebGraphB},
the benefit of the enhanced algorithm (Figure~\ref{fig:sssp2}) is clearly
shown, with a 40\% reduction in execution time compared to the original
SSSP algorith (Figure~\ref{fig:sssp}).
But the Bellman-Ford-style iterations are especially harmful to the MR-MPI and
Hadoop implementations for {WebGraphB}, which required 110 iterations to 
complete; execution times for this data set are show in Table~\ref{t:ssspB}.

@@

In Figure~\ref{f:prbig}, we show the performance 
of the various PageRank
implementations on distributed memory and multi-threaded architectures.
The MR-MPI implementation demonstrated good scalability up to 1024 processors; 
however, it required an order-of-magnitude
more execution time than the matrix-based implementations on Redstorm.  
The distributed memory matrix-based
implementations are competitive with the multi-threaded implementation
in MTGL on the Cray XMT.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_pagerank_big.pdf}
\caption{Scalability comparison of PageRank implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{f:prbig}
\end{figure}

@@

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_cc_big.pdf}
\caption{Scalability comparison of Connected Components implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{f:ccbig}
\end{figure}

>>>>

We ran each of the algorithms on three R-MAT graphs of
different sizes, each on a varying number of processors.  Details of the
input data are shown in Table~\ref{t:rmats}.
The {\it small} problem (around 8M edges) can typically be used
on a single processor without significant out-of-core operations.
The {\it medium} problem (around 134M edges) is a 
problem that can be run
on a single processor with out-of-core operations; larger processor 
configurations, however, do not necessarily need out-of-core operations.
The {\it large} problem (around 2B
edges) requires out-of-core operations and higher processor counts.
%The {\it x-large} problem 
%(around 34B edges) requires most of the machine to run.
All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19, 0.05)$.

@@

The resulting timings give a sense of the inherent scalability of the
MapReduce algorithms as graph size grows on a fixed number of
processors, and of the parallel scalability for computing on a graph
of fixed size on a growing number of processors.  Where available, we
compare the MapReduce algorithm with other parallel implementations,
including more traditional distributed-memory algorithms and 
multi-threaded algorithms
in the Multi-Threaded Graph Library (MTGL)~\cite{MTGL} on the Cray XMT.
We compute parallel efficiency on $p$ processors
as $({time}_{m} \times m) / ({time}_p \times p)$
where $m$ is the smallest number of processors on which the experiment
was run, and ${time}_m$ is the execution time required on $m$ processors.

@@

In Figure~\ref{f:rmat}, we show the scalability of the R-MAT generation 
algorithm (Figure~\ref{fig:rmat2}) for RMAT-20, RMAT-24 and RMAT-28.  For
RMAT-20 and RMAT-24, superlinear speed-up is shown.  This speed-up is due
to the decreased amount of file I/O needed with greater numbers of 
processors; with a larger total memory, a fixed-size problem requires
less I/O with more processors.  For RMAT-28, which requires
significant out-of-core operations, parallel efficiency ranges from 52\% to
97\%.  Comparisons between the enhanced algorithm (Figure~\ref{fig:rmat2}) and
the original algorithm (Figure~\ref{fig:rmat}) showed approximately 
10\% reduction in execution time for the enhanced algorithm.

Also in Figure~\ref{f:rmat}, we compare our MR-MPI implementation with an 
R-MAT generator in MTGL on the Cray XMT.  In the MTGL implementation, 
threads generate edges independently.  A global hash table is used to remove
duplicate edges.  Our MR-MPI implementation runs more quickly than the MTGL
implementation.  The differences in execution time are greatest on 64
processors, where MR-MPI needs the least file I/O for out-of-core operations.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_rmat.pdf}
\caption{Performance of the MR-MPI R-MAT generation algorithm (Figure~\ref{fig:rmat2}) and an MTGL implementation.}
\label{f:rmat}
\end{figure}

\subsection{PageRank results}
\label{subsec:results_pagerank}

In Figure~\ref{f:pr}, we show the performance of the MR-MPI PageRank algorithm
(Figure~\ref{fig:pr2}) compared to a
distributed-memory matrix-based implementation using the linear
algebra toolkit Trilinos~\cite{Trilinos-Overview}.
The matrix-based distributed-memory implementation of PageRank
uses Trilinos' Epetra matrix/vector classes to represent the graph and
PageRank vector.
Rows of matrix $A$ and the associated entries of the PageRank vector $x$
are uniquely assigned to processors; a random permutation of the input
matrix achieves processor load balance.
Interprocessor communication gathers $x$ values for matrix-vector
multiplication and sums partial products into the $y$ vector.
Most communication is point-to-point communication,
but some global communication is needed for computing
residuals and norms of $x$ and $y$.

@@

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_pagerank.pdf}
\caption{Comparison of PageRank implementations using 
MR-MPI (Figure~\ref{fig:pr2}) and 
Trilinos' matrix/vector classes on R-MAT data sets.}
\label{f:pr}
\end{figure}

@@

In Figure~\ref{f:tri}, we show the performance of the triangle finding 
algorithm (Figure~\ref{fig:tri}).  Execution times for this algorithm were
too large to allow the problem to be run with RMAT-28 on our cluster.
Parallel efficiencies for RMAT-20 ranged from 80\% to 140\%; for RMAT-24, they
ranged from 50\% to 120\%.  Comparisons with an MTGL implementation on the 
Cray XMT give mixed results.  When MR-MPI uses the least file I/O for 
out-of-core operations (e.g., RMAT-20 on 32 and 64 processors), it 
outperforms the MTGL implementation.  However, when significant out-of-core
operations are needed (e.g., for RMAT-24), the MTGL implementation is faster.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_tri.pdf}
\caption{Comparison of triangle-finding implementations using
MR-MPI (Figure~\ref{fig:tri}) and MTGL.}
\label{f:tri}
\end{figure}

@@

@@

The execution times for the maximal independent set algorithm
(Figure~\ref{fig:luby}) are shown in Figure~\ref{f:luby}.  Like the R-MAT
generation results, superlinear speed-up of the algorithm occurs for RMAT-20
and RMAT-24, as more of the graph fits into processor memory and less file
I/O is needed.  For RMAT-28, the algorithm requires significant out-of-core
operations. In this case, parallel efficiency is nearly perfect going from 
8 to 64 processors.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_luby.pdf}
\caption{Performance of the MR-MPI maximal independent set algorithm (Figure~\ref{fig:luby}).}
\label{f:luby}
\end{figure}

@@

To compare our MR-MPI implementation with a wider set of implementations,
we performed experiments MR-MPI, Hadoop, PBGL (Parallel
Boost Graph Library)~\cite{PBGL} and a multi-threaded implementation on the
Cray XMT using two web graphs: {WebGraphA} with 13.2M
vertices and 31.9M edges, and {WebGraphB} with 187.6M vertices and 531.9M edges.
In Figure~\ref{f:ssspA}, we show execution times for the SSSP algorithm
using {WebGraphA}.   Like our MR-MPI implementation, the Hadoop 
implementation is a Bellman-Ford-style~\cite{Bellman58,Ford62}  algorithm.  
The XMT and PBGL implementations are based on 
delta-stepping~\cite{MeyerSanders98}, and do not require full iterations
over the entire edge list to advance a breadth-first search.  
We observe that the MR-MPI
implementation runs in less time than the Hadoop implementation, but requires
significantly more time than the XMT and PBGL implementations.  
In experiments with {WebGraphB},
the benefit of the enhanced algorithm (Figure~\ref{fig:sssp2}) is clearly
shown, with a 40\% reduction in execution time compared to the original
SSSP algorithm (Figure~\ref{fig:sssp}).
But the Bellman-Ford-style iterations are especially harmful to the MR-MPI and
Hadoop implementations for {WebGraphB}, which required 110 iterations to 
complete; execution times for this data set are shown in Table~\ref{t:ssspB}.

@@

In Figure~\ref{f:prbig}, we show the performance 
of the various PageRank
implementations on distributed memory and multi-threaded architectures.
The MR-MPI and Trilinos implementations are described in 
Sections~\ref{subsec:graph_pagerank} and~\ref{subsec:results_pagerank},
respectively.
In the multi-threaded MTGL implementation,
rank propagates via adjacency list traversal
in a compressed sparse-row data structure.
To maintain its scalability, code must
be written so that a single thread spawns the loop that processes all
in-neighbors of a given vertex; this detail enables the compiler to generate
hotspot-free code.

The MR-MPI implementation demonstrated good scalability up to 1024 processors; 
however, it required an order-of-magnitude
more execution time than the matrix-based implementations on Redstorm.  
The distributed memory matrix-based
implementations are competitive with the multi-threaded implementation
in MTGL on the Cray XMT.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_pagerank_big.pdf}
\caption{Scalability comparison of PageRank using MapReduce (MR-MPI),
matrix-based (Trilinos), and multi-threaded (MTGL) implementation
on the R-MAT data sets in Table~\ref{t:rmat}.}
\label{f:prbig}
\end{figure}

@@

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_cc_big.pdf}
\caption{Scalability comparison of Connected Components algorithms using 
MapReduce (MR-MPI),
matrix-based/MapReduce hybrid (Trilinos/MR-MPI), and MTGL implementations
on the R-MAT data sets in Table~\ref{t:rmat}.}
\label{f:ccbig}
\end{figure}
